Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-12-19 Ben_Manuscripts/structure_paper/Draft.tex
Modifications based on comments

  %MRS4: are all commercial RO membranes just like this? This seems
  %awfully specific.  
  %MRS4: these two paragraphs are perhaps not
  %ideal.  They describe the process of making the membranes, but that
  %is not what we need as the introduction to THIS paper.  Instead, you
  %need to focus on the strength/issues with RO and NF that motivate this research.

  Commercial RO membranes are thin film composite membranes with a porous
  support layer and an active layer made of a polymer matrix formed by
  interfacial polymerization of a diamine. During the membrane fabrication,
  aqueous diamine solution is allowed to penetrate into the polysulfone support
  layer which is then immersed in trimesoyl chloride, a compound that is
  immiscible in water. Diamine polymerizes with trimesoyl chloride at the
  solution interface and creates a thin film that physically adheres to the
  polysulfone support. The polymer matrix is dense and highly crosslinked
  which leads to low permeability.

  NF membranes are typically made by a phase inversion process. During phase
  inversion, a thin polymer film is precipitated out of a polymer solution.
  Through the use of a non-solvent, a polymer-rich and polymer-poor phase is
  formed. A membrane is created as the polymer-poor phase forms pores in the
  polymer-rich phase. Resultant pores are polydisperse in size which hurts
  membrane selectivity. A second technique used to create NF membranes is call
  track etching in which a polymer film is bombarded with charged particles, then
  chemically etched to create pores. The pores are uniform which benefits
  selectivity, however the membrane has a low porosity and subsequent low
  permeability.

@@

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production.  \cite{feng_scalable_2014} LLC systems created by the monomer 
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied experimentally.
  \cite{smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
  phase (Fig.\ref{fig:assembly}). The presence of ca. 10\% wt water results in the
  H\textsubscript{II} phase. In both cases, monomers assemble into mesophases
  made of hexagonally packed, uniform size, cylinders with hydrophilic head
  groups oriented inward towards the cylinder center.  The hydrophilic region can
  act as a pore for aqueous separations. 
  %MRS4: aren't' there some citations where Gin and noble at least try to separate things (Even at low efficiency) to determine the size? (zhou_supported_2005) that should be cited here on ``act as a pore'' to show these are at least existing (if not usable
 One can envision tailoring the pore
  region for specific separations by changing the monomer chemistry.
%MRS4: can cite the Resel paper paper when changed tails and got different pore-to-pore distances as an example of changing monomer chemistry.

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II} phase LLC membrane studies have been limited primarily to
  Na-GA3C11 with some characterization done after minor structural modifications.
  Resel et al.~varied the length of the monomer tails and the counterion used and
  observed its affect on pore spacing \cite{resel_structural_2000}.

%MRS4: is this sentence needed here? was there an expectation it would? 
%Their study offers no insight into dynamics within the pore. 
In a later study of rejection
  performance, it was shown that membranes formed by Na-GA3C11 cannot perform
  separations of solutes less than 1.2 nm in diameter because the pores are too
  large \cite{zhou_supported_2005}.  We do not yet understand how to controllably
  reduce the effective pore size or how to tune the chemical environment in the
  nanopores for effective water desalination and small organic molecule
  separations. The only source of predictive modeling for LLC systems have been
  macroscopic models which likely do not adequately describe transport at these
  length scales \cite{hatakeyama_water_2011}. It will be challenging to
  efficiently narrow down the large design space in a laboratory setting without
  a robust model.

  %MRS3: this paragraph seems to be a little out of place: should go before a lot of the introductory material, since it lays out what you will be doing in the paper.
  %BJC2: Paragraph moved up here from well below.
  In this study, we build a significantly more realistic atomistic model of LLC
  membranes than, to our knowledge, has ever previously been done, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the model using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure which have been drawn from X-ray diffraction (XRD) experiments
  and in matching ionic conductivity measurements \cite{feng_thin_2016}.

  A molecular level understanding of LLC membrane structure, enabled by
  molecular dynamics simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}),
%MRS4: added as caveat.
  assuming the force fields are sufficiently accurate.
%MRS4: also added.
  with such an atomistic model,
  We can directly observe solute transport and suggest
  governing mechanisms. We can observe how the choice of head group may influence
  pore size for size exclusion driven separations. We can interchange counterions
  which may influence both the pore size and the strength of the Donnan potential
  which affects the degree to which the membrane can exclude charged species.

@@

  %BJC: I think the first sentence of the following paragraph is
  %necessary to put somewhere, but maybe not in this spot exactly
  In order to appropriately model transport, we must first gain a thorough
  understanding of the nanoscopic structure of LLC membranes. Our approach to
  constructing a general model will follow the development of a model of the
  assembly formed by Na-GA3C11 since it has sufficient experimental
  characterization. We have also narrowed our scope to the development of a model
  of the Col\textsubscript{h} phase membrane. Compared to the H\textsubscript{II}
  phase, the Col\textsubscript{h} phase is a simpler starting point, due to the
  absence of water, and has detailed experimental wide-angle X-ray scattering
  (WAXS) patterns useful for reconstructing structural data.  
  %MRS4: you should mention that the two phases appear to have similar pore-to-pore distances, 
  %suggesting they are not very different.

  %MRS4: weird to have a one sentence paragraph.  Maybe combine with next after rephrasing?
  Despite having structural data, there is still information which experiment
  cannot definitively answer. There are several key questions that we will
  investigate.

  Monomers in the Col\textsubscript{h} system are theorized to be partitioned
  into stacked layers which form columnar pores. We %MRS4:want 
  would like to know

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer.!\cite{zhu_methacrylated_2006}.  
  %MRS4: moving citation to end of first reference of a paper. 
%MRS4: rephrase:
However, their estimate 
is based on a simulated
  system containing only 16 total monomers which likely does not sufficiently
  model the chemical environment present in the real system

A separate calculation based on the volume of
  the liquid crystal monomers proposes that there are seven monomers in each
  layer \cite{resel_structural_2000}.  A molecular model orders of magnitude
  larger than any other reported atomistic liquid crystal membrane simulations,
%MR4: added 
as we present here,
  has the best chance of directly answering this question.  We will directly
  change the layer composition and note its effect on membrane structure.

@@

  %BJC: TODO: update colorbar
  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
%MRS3: would be good to have all feature labels agree: either can edit to put the names on, or have a key relating numbers to the names you use.
%BJC2: I will annotate it with the names R-pi, R-alkanes etc. Is that what you mean?
%MRS4: that works fine.  We just need a way to connect what is in the figure with the names used in the text.  Ideally there would not be two sets of labels, just one. 
	\caption{(a) The repeat spacing in the 1D small angle X-ray scattering pattern
		is characteristic of hexagonal packing. The leading peak represents the
		distance between the d\textsubscript{100} planes. Using this distance, we know
		that the distance between pore centers is 4.12 nm. (b) 2D wide angle X-ray
		scattering gives details about repeating features less than 1 nanometer
		apart. Experimentalists have justified each of the 5 major reflections
		present as follows: (I) Aromatic head group $\pi-\pi$ stack 3.7 \AA~apart.
		(II) Monomers arrange vertically in a 2\textsubscript{1} helix. (III) Alkane
 		chain tails pack 4.5 \AA~apart. (IV) Monomer tails tilt with respect to the 
 		membrane plane. (V) As derived from SAXS, pores pack hexagonally and are 
		spaced 4.12 nm apart}
	\label{fig:SWAXS}
 \end{figure}

@@

  Liquid crystal monomers were parameterized using the Generalized AMBER
  Forcefield \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. Atomic charges were assigned using the am1bccsym
  method of molcharge shipped with QUACPAC from Openeye Scientific Software. All
  molecular dynamics simulations were run using Gromacs 2016.
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}
%MRS4: somewhere would need to specify things like time step tau_t, tau_p, etc., so that people could in theory replicate 
the experiments.  Note: we should put scripts and starting configurations (not all that were done but the important ones) in a public repository by the time of submission.  
  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a parameterized
  monomer. Monomers were cooled from 1000K to 50K over 10 nanoseconds.  A low
%MRS3: need to describe this in more detail, but can be in the supplementary. 
%BJC2: would it be best to include the parameterization script
  energy configuration was randomly pulled from the trajectory and charges were
  reassigned using molcharge.
%MRS4: I would give a ``plain text'' version of the script in supplementary.
%MRS3: need to say what molcharge is.
%BJC2: I say in paragraph above : "Atomic charges were assigned using the am1bccsym method of 
%molcharge shipped with QUACPAC from ...". Is that not enough?
%MRS4: sorry, missed that.  Maybe put software command names in \texttt{software} to make it clear what they are.
  Using the new charges, the monomer system was
  annealed again and a random monomer configuration was pulled from the
  trajectory to be used for full system construction (Figure~\ref{fig:python}a).
  See supplemental information for further detail.

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure when a
  reasonable value is chosen (See Supplemental Information). The pore radius is chosen to be
  0.6 nm in our initial configurations because the pore size is estimated to be
  $\approx$ 1.2 nm.
%MRS4: provide citation here for 1.2 number.
 The initial distance between pores also has little effect on
  the the equilibrated structure. However, one should not start them too close or
  there will be high energy repulsions during early equilibration. We chose an
  initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental
  value of 4.12 nm.  The distance between layers, the rotation of the layers with
  respect to adjacent layers, and the number of monomers per layer do influence
  the equilibrium structure and require further justification for their choices.
  We rely on experimental data to inform them.

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was rotated
  so the plane of the aromatic head groups would be coplanar with the xy plane.
  We explore two different initial layer spacings. The first is exactly equal to
  R-$\pi$ with layers placed so aromatic rings are stacked 3.7 \angstrom~apart
  in the z-direction. A second system is explored with an initial layer spacing
  of 5 \angstrom. A third system with an initial layer spacing of 10 \angstrom~was
  %MRS4: this is probably OK here, but you do want to move away from describing things historically (the order in which the ideas were had and tested) and the logical ordering to communicate the story.  You want to do the second whenever possible. Think about this going over the paper and try to shift from historical presentation of results to story-based ordering.
  briefly explored. When layers are spaced out too far, they will collapse
  on each other while simultaneously slipping in between layers of adjacent pores,
  which leads to an artificially thick membrane with pores spaced closely
  together. 
  Further details of simulations with large layer spacing are in the supplemental information.
%There will be no further discussion of this system but the interested
%  reader can learn more about it in the supplemental information.

@@

%MRS4: describe what the points are, people don't want to go back and look.
  As outlined in (\ref{point:monomernum}) the number of monomers in each layer is
  unknown. We tested configurations constructed with a varied number of
  monomers per layer. Systems were built in the offset and parallel displaced
  configurations with 4, 5, 6, 7 and 8 monomers per layer.

@@

  % BJC: Need to make it clear that this was not done on all systems

  In order to fully match synthetic procedures, we created a crosslinking
  algorithm that can be applied to equilibrated structures. The purpose of 
  crosslinking
  is to maintain macroscopic alignment of the crystalline domains, 
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data. The
  algorithm was developed based on the known reaction mechanism. Crosslinking of
  this system is a free radical polymerization (FRP) taking place between
  terminal vinyl groups present on each of the three monomer tails. FRPs require
  an initiator which bonds to the system, meaning new atoms are introduced into
  the system. For simplicity, the initiator was simulated as hydrogen and made
  present in the simulation by including them in all possible locations where an 
  addition could occur as dummy atoms. The crosslinking procedure is carried out
  iteratively. During each iteration, bonding carbon atoms are chosen based on
  a distance cut-off. The topology is updated with new bonds and dummy hydrogen
  atoms are changed to appropriate hydrogen types. Head-to-tail addition was the
  only propagation mode considered due to its dominance in the real system.
%MRS3: citation why it dominates the real system.
%BJC2: Yea. There are a few citations that I need to make. I know they exist
% but I need to track them down. I've prioritized that last. 
  Direction of attack was not considered because the resultant mixture is
  racemic.

@@

  The colorbars on all diffraction patterns are normalized relative to
  R-alkanes. We believe that the alkane-alkane density, averaged over all angles,
  is the feature most likely to be replicated between experiment and simulation.
  Other features are dependent on system ordering which is likely to have some
  dependence on initial configuration. We calculated the average intensity within
  R-alkanes of the experimental pattern, $I_{avg}$.  We exclude intensities
  within $\pm$ 30\degree~of the $q_r=0$ axis, since the simulated patterns differ
  from experiment in those regions. 
%MRS4: added to connect.
Specifically, in contrast to the experimental WAXS pattern,
  R-$\pi$ appearing in simulated diffraction patterns intersects with R-alkanes
  (See Fig. \ref{fig:XRDsim}). We multiplied $I_{avg}$ by a scaling factor of
  2.5. Intensities that appear in the experimental pattern $\geq$ 2.5$I_{avg}$
  receive colorbar values of 1. We apply the same scaling method to the simulated
  patterns. We carefully chose a scaling factor of 2.5 in order to (1) visibly
  display all features in the experimental pattern and (2) to allow us to compare
  the relative intensities of features between simulated and experimental
  diffraction patterns.

  % BJC: I think this figure belongs somewhere else 
%MRS4: yeah, probably. Possibly supplemental information.
  \begin{figure}
	\centering
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=\textwidth]{sandwiched.png}
		\caption{}\label{fig:sandwiched}
	\end{subfigure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=\textwidth]{PD.png}
		\caption{}\label{fig:pd}
	\end{subfigure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Tshaped.png}
		\caption{}\label{fig:tshaped}
	\end{subfigure}
	\vskip\baselineskip
	\begin{subfigure}[b]{0.475\textwidth}
		\centering
		\includegraphics[width=\textwidth]{sandwichedlayers.png}
		\caption{}\label{fig:sandwichedlayers}
	\end{subfigure}
	\begin{subfigure}[b]{0.475\textwidth}
		\centering
		\includegraphics[width=\textwidth]{offsetlayers.png}
		\caption{}\label{fig:offsetlayers}
	\end{subfigure}
	\caption{(a) Sandwiched benzene dimers stack 3.8 \angstrom~apart. (b) Parallel-Displaced benzene dimers stack
	3.4 \angstrom~vertically and 1.6 \angstrom~horizontally apart. (c) T-shaped benzene dimers stack 5.0 \angstrom~apart. 
	(d) Two monomer layers stacked in the sandwiched configuration (e) Two monomer layers stacked in the parallel-displaced
	configuration }\label{fig:stacking}
  \end{figure}

@@

  \begin{figure}
	\centering
	\includegraphics[width=\linewidth]{p2p.png}
	\caption{Systems built with 5 monomers per layer in a parallel
		displaced configuration result in a pore spacing closest to the experimental
		value of 4.12 nm. The pore spacing of the model increases as number of monomers
		in each layer increases. The pore spacing of a system starting in the
		sandwiched configuration is systematically 
%MRS4: added
slightly
lower than that started in an offset
		configuration. }~\label{fig:p2p}
  \end{figure}

@@

%MRS4: can we come up with another name?  I still think disordered pores is a decent name, since there is no pi-stacking layering.  ``structures that form when layers are 5 AA apart'' is not so good.  or ``more disordered''?
%  The $z$-direction correlation functions show that layers in our model prefer
%  to stack further apart than 3.7 \angstrom~as suggested by experiment.
  When systems are built with layers stacked 5.0 \AA~apart then
  equilibrated, we observe long-term stability of a qualitatively different
  configuration, suggesting that we have found another metastable free energy
  basin, further corroborating (\ref{point:metastable}). We studied this type of
  system in both the parallel displaced and sandwiched configurations.

>>>>

  %MRS4: are all commercial RO membranes just like this? This seems
  %awfully specific. 
  %BJC3: Yea, pretty much all commercial membranes are TFC polyamide
  % but I edited it to make it less specific.

  %MRS4: these two paragraphs are perhaps not ideal.  They describe the process
  %of making the membranes, but that is not what we need as the introduction to
  %THIS paper.  Instead, you need to focus on the strength/issues with RO and
  %NF that motivate this research.
  % BJC3: The point I am trying to make is that the way the membranes are made 
  % results in either low selectivity or low permeability. I could just state
  % that, or give a little background on the processes to make the point more
  % tangible

  Commercial RO membranes are thin film composite membranes with a porous
  support layer and an active layer made of a polymer matrix formed by
  interfacial polymerization. During synthesis, a step-growth polymerization
  reaction occurs at the interface of an aqueous monomer solution, which has been
  soaked into the support layer, and an organic solution of a second monomer. The
  polymer matrix is dense and highly crosslinked which gives high rejections but
  low permeability.

  %During membrane fabrication,
  %aqueous diamine solution is allowed to penetrate into the polysulfone support
  %layer which is then immersed in trimesoyl chloride, a compound that is
  %immiscible in water. Diamine polymerizes with trimesoyl chloride at the
  %solution interface and creates a thin film that physically adheres to the
  %polysulfone support.

%  solution.  Through the use of a non-solvent, a polymer-rich and polymer-poor
%  phase is formed. A membrane is created as the polymer-poor phase forms pores in
%  the polymer-rich phase. 
%  During phase inversion, a thin polymer film is precipitated out of a polymer
  NF membranes are typically made by a phase inversion process. The most widely
  used phase inversion process is immersion precipitation, during which a polymer
  dissolved in a solvent is submerged in a non-solvent. A solid, porous membrane 
  is all that remains once all solvent has been removed by non-solvent exchange. 
  Resultant pores are polydisperse in size which hurts membrane selectivity.  A
  second technique used to create NF membranes is call track etching in which a
  polymer film is bombarded with charged particles, then chemically etched to
  create pores. The pores are uniform which benefits selectivity, however the
  membranes have a low porosity and subsequent low permeability.

@@

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production.  \cite{feng_scalable_2014} LLC systems created by the monomer
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied
  experimentally.
  \cite{smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h} phase
  (Fig.\ref{fig:assembly}). The presence of ca. 10\% wt water results in the
  H\textsubscript{II} phase. In both cases, monomers assemble into mesophases
  made of hexagonally packed, uniform size, cylinders with hydrophilic head
  groups oriented inward towards the cylinder center.  The hydrophilic region can
  act as a pore for aqueous separations \cite{zhou_supported_2005}.  One can
  envision tailoring the pore region for specific separations by changing the
  monomer chemistry \cite{resel_h2-phase_2000}.

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II} phase LLC membrane studies have been limited primarily to
  Na-GA3C11 with some characterization done after minor structural modifications.
  Resel et al.~varied the length of the monomer tails and the counterion used and
  observed its affect on pore spacing \cite{resel_structural_2000}.  In a later
  study of rejection performance, it was shown that membranes formed by Na-GA3C11
  cannot perform separations of solutes less than 1.2 nm in diameter because the
  pores are too large \cite{zhou_supported_2005}.  We do not yet understand how
  to controllably reduce the effective pore size or how to tune the chemical
  environment in the nanopores for effective water desalination and small organic
  molecule separations. The only source of predictive modeling for LLC systems
  have been macroscopic models which likely do not adequately describe transport
  at these length scales \cite{hatakeyama_water_2011}. It will be challenging to
  efficiently narrow down the large design space in a laboratory setting without
  a robust model.

  In this study, we build a significantly more realistic atomistic model of LLC
  membranes than, to our knowledge, has ever previously been done, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the model using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure which have been drawn from X-ray diffraction (XRD) experiments
  and in matching ionic conductivity measurements \cite{feng_thin_2016}.

  A molecular level understanding of LLC membrane structure, enabled by
  molecular dynamics simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  with such an atomistic model, We can directly observe solute transport and
  suggest governing mechanisms. We can observe how the choice of head group may
  influence pore size for size exclusion driven separations. We can interchange
  counterions which may influence both the pore size and the strength of the
  Donnan potential which affects the degree to which the membrane can exclude
  charged species.

@@

  %BJC: I think the first sentence of the following paragraph is
  %necessary to put somewhere, but maybe not in this spot exactly
  In order to appropriately model transport, we must first gain a thorough
  understanding of the nanoscopic structure of LLC membranes. Our approach to
  constructing a general model will follow the development of a model of the
  assembly formed by Na-GA3C11 since it has sufficient experimental
  characterization. We have also narrowed our scope to the development of a model
  of the Col\textsubscript{h} phase membrane. Compared to the H\textsubscript{II}
  phase, the Col\textsubscript{h} phase is a simpler starting point, due to the
  absence of water, and has detailed experimental wide-angle X-ray scattering
  (WAXS) patterns useful for reconstructing structural data. The two
  phases appear to share similar structural characteristics since the pore 
  spacings in each system are in close agreement \cite{resel_structural_2000,feng_thin_2016}.

  Despite having structural data, there is still information which experiment
  cannot definitively answer. Our current level of understanding suggests that
  monomers partition into layers with head groups facing the center, then 
  stack on top of each other to create pores that pack together hexagonally.
  There are several key questions that we will investigate in order to
  expand our understanding.

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer \cite{zhu_methacrylated_2006}. Their 
  model likely does not sufficiently model the chemical environment present
  in the real system since it only contains 16 total monomers.
%BJC3: rephrased
%  However, their estimate is based on a simulated system that contains 16 total
%  monomers which likely does not sufficiently model the chemical environment
%  present in the real system
  A separate calculation based on the volume of the liquid crystal monomers
  proposes that there are seven monomers in each layer
  \cite{resel_structural_2000}.  A molecular model orders of magnitude larger
  than any other reported atomistic liquid crystal membrane simulations, as we
  present here, has the best chance of directly answering this question.  We will
  directly change the layer composition and note its effect on membrane
  structure.

@@

  %BJC: TODO: update colorbar
  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
%MRS3: would be good to have all feature labels agree: either can edit to put the names on, or have a key relating numbers to the names you use.
%BJC2: I will annotate it with the names R-pi, R-alkanes etc. Is that what you mean?
%MRS4: that works fine.  We just need a way to connect what is in the figure with the names used in the text.  Ideally there would not be two sets of labels, just one.
%BJC3: Got it, will do once I have more time to focus on figures 
	\caption{(a) The repeat spacing in the 1D small angle X-ray scattering pattern
		is characteristic of hexagonal packing. The leading peak represents the
		distance between the d\textsubscript{100} planes. Using this distance, we know
		that the distance between pore centers is 4.12 nm. (b) 2D wide angle X-ray
		scattering gives details about repeating features less than 1 nanometer
		apart. Experimentalists have justified each of the 5 major reflections
		present as follows: (I) Aromatic head group $\pi-\pi$ stack 3.7 \AA~apart.
		(II) Monomers arrange vertically in a 2\textsubscript{1} helix. (III) Alkane
 		chain tails pack 4.5 \AA~apart. (IV) Monomer tails tilt with respect to the 
 		membrane plane. (V) As derived from SAXS, pores pack hexagonally and are 
		spaced 4.12 nm apart}
	\label{fig:SWAXS}
 \end{figure}

@@

  Liquid crystal monomers were parameterized using the Generalized AMBER
  Forcefield \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. Atomic charges were assigned using the am1bccsym
  method of \texttt{molcharge} shipped with QUACPAC from Openeye Scientific
  Software. All molecular dynamics simulations were run using Gromacs 2016.
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}
  the experiments.  An ensemble of characteristic, low-energy vacuum monomer
  configurations were constructed by applying a simulated annealing process to a
  parameterized monomer. Monomers were cooled from 1000K to 50K over 10
  nanoseconds.  A low energy configuration was randomly pulled from the
  trajectory and charges were reassigned using \texttt{molcharge}.  Using the new
  charges, the monomer system was annealed again and a random monomer
  configuration was pulled from the trajectory to be used for full system
  construction (Figure~\ref{fig:python}a).  See supplemental information for
  further detail.

% Note: we should put scripts and starting configurations (not all that were
% done but the important ones) in a public repository by the time of
% submission. 
%MRS4: I would give a ``plain text'' version of the script in supplementary. BJC: referring to parameterization script
%MRS4: somewhere would need to specify things like time step tau_t, tau_p, etc., so that people could in theory replicate

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure when a
  reasonable value is chosen (See Supplemental Information). The pore radius is
  chosen to be 0.6 nm in our initial configurations because the pore size is
  estimated to be $\approx$ 1.2 nm \cite{zhou_supported_2005}.  The initial
  distance between pores also has little effect on the the equilibrated
  structure. However, one should not start them too close or there will be high
  energy repulsions during early equilibration. We chose an initial pore spacing
  of 4.5 nm, $\sim$10\% larger than the experimental value of 4.12 nm.  The
  distance between layers, the rotation of the layers with respect to adjacent
  layers, and the number of monomers per layer do influence the equilibrium
  structure and require further justification for their choices.  We rely on
  experimental data to inform them.

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was
  rotated so the plane of the aromatic head groups would be coplanar with the xy
  plane.  We explore two different initial layer spacings. The first is exactly
  equal to R-$\pi$ with layers placed so aromatic rings are stacked 3.7
  \angstrom~apart in the z-direction. A second system is explored with an initial
  layer spacing of 5 \angstrom. A third system with an initial layer spacing of
  10 \angstrom~was briefly explored. When layers are spaced out too far, they
  will collapse on each other while simultaneously slipping in between layers of
  adjacent pores, which leads to an artificially thick membrane with pores spaced
  closely together.  Further details of simulations with large layer spacing are
  in the supplemental information.

@@

  The number of monomers in each layer is unknown, as state in
  (\ref{point:monomernum}). We tested configurations constructed with a varied
  number of monomers per layer. Systems were built in the offset and parallel
  displaced configurations with 4, 5, 6, 7 and 8 monomers per layer.

@@

  In order to fully match synthetic procedures, we created a crosslinking
  algorithm that can be applied to equilibrated structures. The purpose of
  crosslinking is to maintain macroscopic alignment of the crystalline domains,
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data. The
  algorithm was developed based on the known reaction mechanism. Crosslinking of
  this system is a free radical polymerization (FRP) taking place between
  terminal vinyl groups present on each of the three monomer tails. FRPs require
  an initiator which bonds to the system, meaning new atoms are introduced into
  the system. For simplicity, the initiator was simulated as hydrogen and made
  present in the simulation by including them in all possible locations where an
  addition could occur as dummy atoms. The crosslinking procedure is carried out
  iteratively. During each iteration, bonding carbon atoms are chosen based on a
  distance cut-off. The topology is updated with new bonds and dummy hydrogen
  atoms are changed to appropriate hydrogen types. Head-to-tail addition was the
  only propagation mode considered due to its dominance in the real system.
  Direction of attack was not considered because the resultant mixture is
  racemic.

@@

  The colorbars on all diffraction patterns are normalized relative to
  R-alkanes. We believe that the alkane-alkane density, averaged over all angles,
  is the feature most likely to be replicated between experiment and simulation.
  Other features are dependent on system ordering which is likely to have some
  dependence on initial configuration. We calculated the average intensity within
  R-alkanes of the experimental pattern, $I_{avg}$.  We exclude intensities
  within $\pm$ 30\degree~of the $q_r=0$ axis, since the simulated patterns differ
  from experiment in those regions.  Specifically, in contrast to the
  experimental WAXS pattern, R-$\pi$ appearing in simulated diffraction patterns
  intersects with R-alkanes (See Fig. \ref{fig:XRDsim}). We multiplied $I_{avg}$
  by a scaling factor of 2.5. Intensities that appear in the experimental pattern
  $\geq$ 2.5$I_{avg}$ receive colorbar values of 1. We apply the same scaling
  method to the simulated patterns. We carefully chose a scaling factor of 2.5 in
  order to (1) visibly display all features in the experimental pattern and (2)
  to allow us to compare the relative intensities of features between simulated
  and experimental diffraction patterns.

  % BJC: I think this figure belongs somewhere else 
%MRS4: yeah, probably. Possibly supplemental information.
%  \begin{figure}
%	\centering
%	\begin{subfigure}[b]{0.32\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{sandwiched.png}
%		\caption{}\label{fig:sandwiched}
%	\end{subfigure}
%	\begin{subfigure}[b]{0.32\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{PD.png}
%		\caption{}\label{fig:pd}
%	\end{subfigure}
%	\begin{subfigure}[b]{0.32\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{Tshaped.png}
%		\caption{}\label{fig:tshaped}
%	\end{subfigure}
%	\vskip\baselineskip
%	\begin{subfigure}[b]{0.475\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{sandwichedlayers.png}
%		\caption{}\label{fig:sandwichedlayers}
%	\end{subfigure}
%	\begin{subfigure}[b]{0.475\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{offsetlayers.png}
%		\caption{}\label{fig:offsetlayers}
%	\end{subfigure}
%	\caption{(a) Sandwiched benzene dimers stack 3.8 \angstrom~apart. (b) Parallel-Displaced benzene dimers stack
%	3.4 \angstrom~vertically and 1.6 \angstrom~horizontally apart. (c) T-shaped benzene dimers stack 5.0 \angstrom~apart. 
%	(d) Two monomer layers stacked in the sandwiched configuration (e) Two monomer layers stacked in the parallel-displaced
%	configuration }\label{fig:stacking}
%  \end{figure}

@@

  \begin{figure}
	\centering
	\includegraphics[width=\linewidth]{p2p.png}
	\caption{Systems built with 5 monomers per layer in a parallel
		displaced configuration result in a pore spacing closest to the experimental
		value of 4.12 nm. The pore spacing of the model increases as number of monomers
		in each layer increases. The pore spacing of a system starting in the
		sandwiched configuration is systematically slightly lower than that started in
		an offset configuration. }~\label{fig:p2p}
  \end{figure}

@@

%MRS4: can we come up with another name?  I still think disordered pores is a decent name, since there is no pi-stacking layering.  ``structures that form when layers are 5 AA apart'' is not so good.  or ``more disordered''?
%BJC3: agreed, I'll stick with disordered pores
%  The $z$-direction correlation functions show that layers in our model prefer
%  to stack further apart than 3.7 \angstrom~as suggested by experiment.
  When systems are built with layers stacked 5.0 \AA~apart then equilibrated,
  we observe long-term stability of a qualitatively different configuration,
  suggesting that we have found another metastable free energy basin, further
  corroborating (\ref{point:metastable}). For reasons to become aparent later,
  we will refer to this class of systems as the disordered pore basin. 
  Systems with layers initially stacked 3.7 \AA~apart will be referred to
  as the ordered basin. We studied each basin in both the parallel
  displaced and sandwiched configurations.
